Model Recognition as Planning

### 2018-09-21 doc/icaps-19/model-recognition.tex
paper icaps19: pages [1-4] stable. Ready for review

@@

\subsection{The \strips\ edit distance}
We define two edit \emph{operations} on a schema $\xi\in\mathcal{M}$ that belongs to a \strips\ action model $\mathcal{M}\in M$:
\begin{itemize}
\item {\em Deletion}. An element in ${\mathcal I}_{\Psi,\xi}$ is removed from any of these three lists $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the operator schema $\xi\in\mathcal{M}$ such that the resulting schema is a {\em well-defined} \strips\ action schema.
\item {\em Insertion}. An element in ${\mathcal I}_{\Psi,\xi}$ is added to any of these three lists $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the operator schema $\xi\in\mathcal{M}$ s.t. the resulting schema is {\em well-defined}.
\end{itemize}

@@

The observation of plan executions that are generated with $\xi$ action schema constrain further the space of possible schema. In this case we talk about {\em semantic constraints} that can also be added the $\mathcal{C}$ set. We define now an edit distance to asses the matching of a \strips\ model with respect to an observation of a plan execution.

@@

Remarkably, the {\em observation edit distance} allows us to elicit the likelihood of the observations given a candidate model. It can be argued that the shorter this  distance the better the given model explains the given observation and hence, the higher the likelihood $P(\tau|\mathcal{M})$. In particular this distance is maximum when it requires fully editing all the schemas in the model while it is minimum when the given model can produce the input observation without introducing any change. Intermmediate distance values reflect how far are the models from explaining the input observations. In this work we map the {\em observation edit distance} into a likelihood with the following expression $P(\tau|\mathcal{M}) = 1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$.

@@

\subsection{Conclusions}
\label{sec:conclussions}
The task of {\em model recognition} can also be understood as a classification task where each class is represented with a different planning model and the observed plan execution is the example to classify. With this regard, planning model that is associated to a class is acting as a class prototype that the summarizes all the plan executions that could be synthesized with that model or in other words, all the examples that belong to that class.

In this work we do not assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}. A related approach is recently followed for {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the PDDL models of two different agents.

>>>>

The task of {\em model recognition} can also be understood as a classification task where each class is represented with a different planning model and the observed plan execution is the example to classify. In this case, the planning model that is associated to a class is acting as a class prototype that the summarizes all the plan executions that could be synthesized with that model or in other words, all the examples that belong to that class.

@@

\subsection{Edit distances for the \strips\ planning model}
We define two edit \emph{operations} on a schema $\xi\in\mathcal{M}$ that belongs to a \strips\ action model $\mathcal{M}\in M$:
\begin{itemize}
\item {\em Deletion}. An element is removed from any of these three lists $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the operator schema $\xi\in\mathcal{M}$ such that the resulting schema is a {\em well-defined} \strips\ action schema.
\item {\em Insertion}. An element in ${\mathcal I}_{\Psi,\xi}$ is added to any of these three lists $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the operator schema $\xi\in\mathcal{M}$ s.t. the resulting schema is {\em well-defined}.
\end{itemize}

@@

The observation of plan executions, generated with an action model $\mathcal{M}$, is {\em semantic knowledge} that constrain further the space of possible schema $\xi\in \mathcal{M}$. In this case, we talk about {\em observation constraints} that can also be added to the $\mathcal{C}$ set. In addition, this new kind of model constraints allow us to define an edit distance to asses the matching of a \strips\ model with respect to an observation of a plan execution.

@@

Remarkably, the {\em observation edit distance} allows us to elicit the likelihood of the observations given a candidate model. It can be argued that the shorter this  distance the better the given model explains the given observation and hence, the higher the $P(\tau|\mathcal{M})$ likelihood. In particular this distance is maximum when it requires fully editing all the schemas in the model while it is minimum when the given model is already able to produce the input observation without introducing any change. Intermmediate distance values reflect how far models are from explaining the input observations. In this work we map the {\em observation edit distance} into a $P(\tau|\mathcal{M})$ likelihood with the following expression, $P(\tau|\mathcal{M}) = 1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$.

@@

\subsection{Conclusions}
\label{sec:conclussions}
In this work we do not assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}. A related approach is recently followed for {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the PDDL models of two different agents.
